---
title: "Assessment of a Potential Velocity Barrier for Adult Steelhead Migration on the South Fork Clearwater River - PRELIMINARY"
author:
- Michael W. Ackerman:
    email: mikea@nezperce.org
    institute: npt_mccall
    correspondence: yes
- Ryan N. Kinzer:
    email: ryank@nezperce.org
    institute: npt_mccall
    correspondence: no
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    theme: yeti
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: yes
      smooth_scroll: yes
    number_sections: yes
    pandoc_args:
    - "--lua-filter=../templates/scholarly-metadata.lua"
    - "--lua-filter=../templates/author-info-blocks.lua"
    - "--lua-filter=../templates/pagebreak.lua"
institute:
- npt_mccall: Nez Perce Tribe, Department of Fisheries Resources Management, 14054 Burr Dr., McCall, Idaho, 83638, USA
csl: "../templates/journal-of-archaeological-science.csl"
bibliography: AckermanLibrary.bib
always_allow_html: yes
---

<!-- the following inserts MHE logo into header -->

```{=html}
<script>
   $(document).ready(function() {
     $head = $('#header');
     $head.prepend('<img src=\"DFRM.png\" style=\"float: right;width: 150px;\"/>')
   });
</script>

<style>
p.caption {
  font-size: 0.8em;
}
</style>
```

```{r setup, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 600
)

options(knitr.kable.NA = '--')

```

```{r load-libraries}
# run this if running chunks directly

# for formatting
# library(kableExtra)
# library(ggpubr)

# for analysis

```



# Background

A section of the South Fork Clearwater River near milepost (MP) 28 has been considered a potential barrier to adult steelhead migration due to high velocities, at least during particular flows. The high velocities are hypothesized to hinder or prevent some adult steelhead from migrating upstream of the barrier, limiting their use of available spawning habitat upstream of the high-velocity section in the mainstem South Fork Clearwater River and tributaries. Namely, previous work has indicated that stream velocities within the section likely impede upstream passage for adult steelhead at flows greater than approximately 1,000 cfs (~600 cfs at the previous Elk City gage) [@Timm2017; @Timm2018].



## Study area

Consider a study area map with locations of arrays (SC1, SC2, SC3, SC4, CRA), location of potential velocity barrier (milepost 28, rm 44, rkm 71), and streams with juvenile steelhead release groups (mainstem, Clear Creek, Meadown Creek, Newsome Creek).

```{r load-data}

```

# Assessment

## Spawn Year and Release Group

## Fish Length

## Ocean Age

## Water Level

# Discussion

First Newsome Creek juvenile releases occurred in spring 2022; 1-ocean adults will return to SF Clearwater in spring 2024 and 2-ocean adults in spring 2025.


# Literature Cited

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

