---
title: "Assessment of a Potential Velocity Barrier for Adult Steelhead Migration on the South Fork Clearwater River - PRELIMINARY"
author:
- Michael W. Ackerman:
    email: mikea@nezperce.org
    institute: npt_mccall
    correspondence: yes
- Ryan N. Kinzer:
    email: ryank@nezperce.org
    institute: npt_mccall
    correspondence: no
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    theme: yeti
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: yes
      smooth_scroll: yes
    number_sections: yes
    pandoc_args:
    - "--lua-filter=../templates/scholarly-metadata.lua"
    - "--lua-filter=../templates/author-info-blocks.lua"
    - "--lua-filter=../templates/pagebreak.lua"
institute:
- npt_mccall: Nez Perce Tribe, Department of Fisheries Resources Management, 14054 Burr Dr., McCall, Idaho, 83638, USA
csl: "../templates/journal-of-archaeological-science.csl"
bibliography: AckermanLibrary.bib
always_allow_html: yes
---

<!-- the following inserts MHE logo into header -->

```{=html}
<script>
   $(document).ready(function() {
     $head = $('#header');
     $head.prepend('<img src=\"DFRM.png\" style=\"float: right;width: 150px;\"/>')
   });
</script>

<style>
p.caption {
  font-size: 0.8em;
}
</style>
```

```{r setup, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 600
)

options(knitr.kable.NA = '--')

```

```{r load-libraries}
# run this if running chunks directly

# for formatting
# library(kableExtra)
# library(ggpubr)

# for analysis

```



# Background

Some background text here...

## Study area

Consider a study area map with locations of arrays (SC1, SC2, SC3, SC4, CRA), location of potential velocity barrier (milepost 28, rm 44, rkm 71), and streams with juvenile steelhead release groups (mainstem, Clear Creek, Meadown Creek, Newsome Creek).

```{r load-data}

```

# Assessment

## Spawn Year and Release Group

## Fish Length

## Ocean Age

## Water Level

# Discussion

First Newsome Creek juvenile releases occurred in spring 2022; 1-ocean adults will return to SF Clearwater in spring 2024 and 2-ocean adults in spring 2025.


# Literature Cited

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

