---
title: "Assessment of a Potential Velocity Barrier for Adult Steelhead Migration on the South Fork Clearwater River - PRELIMINARY"
author:
- Michael W. Ackerman:
    email: mikea@nezperce.org
    institute: npt_mccall
    correspondence: yes
- Ryan N. Kinzer:
    email: ryank@nezperce.org
    institute: npt_mccall
    correspondence: no
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    theme: yeti
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: yes
      smooth_scroll: yes
    number_sections: yes
    pandoc_args:
    - "--lua-filter=../templates/scholarly-metadata.lua"
    - "--lua-filter=../templates/author-info-blocks.lua"
    - "--lua-filter=../templates/pagebreak.lua"
institute:
- npt_mccall: Nez Perce Tribe, Department of Fisheries Resources Management, 14054 Burr Dr., McCall, Idaho, 83638, USA
csl: "../templates/journal-of-archaeological-science.csl"
bibliography: AckermanLibrary.bib
always_allow_html: yes
---

<!-- the following inserts the NPT logo into header -->
```{=html}
<script>
   $(document).ready(function() {
     $head = $('#header');
     $head.prepend('<img src=\"DFRM.png\" style=\"float: right;width: 150px;\"/>')
   });
</script>

<style>
p.caption {
  font-size: 0.8em;
}
</style>
```

```{r setup, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 600
)

options(knitr.kable.NA = '--')

```

```{r load-libraries}
# run this if running chunks directly

# for formatting
library(kableExtra)

# for analysis
library(tidyverse)
library(PITcleanr)
library(here)
library(sf)
library(ggplot2)
library(janitor)

```


# Background

A section of the South Fork Clearwater River near milepost (MP) 28 has been considered a potential barrier to adult steelhead migration due to high velocities, at least during particular flows. These high water velocities are hypothesized to hinder, or even prevent, some adult steelhead from migrating upstream of the barrier, limiting their use of available spawning habitat upstream of the high-velocity section in the mainsteam South Fork Clearwater River and tributaries. Namely, previous work has indicated that stream velocities within the section likely impede upstream passage for adult steelhead at flows greater than approximately 1,000 cfs (~600 cfs at the previous Elk City gage) [@Timm2017; @Timm2018].

This document provides a preliminary assessment of adult steelhead passage at the putative velocity barrier (hereafter barrier) based on detections of PIT-tagged adults in the South Fork Clearwater River, including at sites SC3 and SC4 located downstream and upstream of the barrier, respectively. In particular, we're interested in whether PIT-tag observations suggest that adults passing SC3 "convert" to locatons upstream of the barrier (e.g., SC4) at rates lower than expected, suggesting that the barrier hinders or prevents passage at least for some adults. Although the evaluation is focused on steelhead, we include data from PIT-tagged adult Chinook salmon for comparison.
<!--Continue, if necessary-->

```{r load-data}
load(here("data/derived_data/config.rda"))
load(here("data/derived_data/sf_clearwater_passage_data.rda"))
load(here("data/derived_data/sf_clearwater_mean_daily_cfs.rda")) ; rm(sf_elk_daily_cfs, sf_elk_gage_info, sf_stites_gage_info)

# environmental data from the SC4 probe
sc4_env = read_csv(here("data/raw_data/SC4_enviro_5f1ff362-1201-4136-a630-0ba05c141177.csv")) %>%
  rename(node = slug...1,
         metric = slug...2) %>%
  mutate(date = date(read_at)) %>%
  group_by(node, metric, date) %>%
  summarise(mean = round(mean(value), 2)) %>%
  ungroup()

```

```{r data-prep}
# detection by tag in wide format
comp_filter_wide = comp_filter %>%
  select(species,
         spawn_year,
         tag_code,
         node,
         min_det,
         travel_time) %>%
  # for instances where a tag has two detections at a single site, get first detection
  group_by(species, spawn_year, tag_code, node) %>%
  slice(which.min(min_det)) %>%
  # need to convert datetime columns to character before pivot_wider
  mutate(min_det = as.character(min_det),
         travel_time = as.character(travel_time)) %>%
  # pivot wider
  group_by(species, spawn_year, tag_code) %>%
  pivot_wider(names_from = node,
              values_from = c(min_det, travel_time)) %>%
  ungroup()

# prepare data frame for analysis  
sf_df = rbind(sy2022_chnk_ch,
              sy2023_chnk_ch,
              sy2022_sthd_ch,
              sy2023_sthd_ch) %>%
  select(species, spawn_year, tag_code, cap_hist) %>%
  left_join(comp_filter_wide) %>%
  # attach some useful information from LGRTrappindDB
  left_join(sf_lgr_df %>%
              select(-spawn_year)) %>%
  # attach some useful mark and release information from DART
  left_join(sf_dart_obs %>%
              select(tag_id,
                     mark_site,
                     mark_date,
                     rel_site,
                     rel_date) %>%
              distinct(),
            by = c("tag_code" = "tag_id")) %>%
  # join water level data from the SC4 probe, using date when fish was last detected at SC1, SC2, or SC3
  mutate(tmp = date(pmax(min_det_SC1, min_det_SC2, min_det_SC3, na.rm = T))) %>%
  left_join(sc4_env %>%
              filter(metric == "water_level") %>%
              select(date,
                     sc4_water_level_m = mean),
            by = c("tmp" = "date")) %>%
  # join water level data from stites usgs gage 13338500
  left_join(sf_stites_daily_cfs %>%
              select(Date,
                     daily_mean_cfs),
            by = c("tmp" = "Date")) %>%
  select(-tmp) %>%
  # did fish migrate past the potential velocity barrier?
  mutate(pass_sc3 = !is.na(min_det_SC3),
         success = !is.na(min_det_SC4) | !is.na(min_det_CRA)) %>%
  select(species,
         spawn_year,
         tag_code,
         cap_hist,
         pass_sc3,
         success,
         everything())

```

## Study area

The section of the river considered to be a potential velocity barrier is located on river kilometer (rkm) 71 of the South Fork Clearwater River (Figure \@ref(fig:study-map)). Figure \@ref(fig:study-map) and Table \@ref(tab:sites-table) additionally show each of the instream PIT tag detection sites (IPTDS) used in the evaluation to detect adult steelhead and Chinook salmon.

```{r study-map, message = F, fig.cap = "Map of the study area including the locations of each instream PIT tag detection system and the approximate location of the potential velocity barrier shown in red."}
# sites of interest
sf_sites = c("SC1",  # rkm 1
             "SC2",  # rkm 2   
             "SC3",  # rkm 60
             "SC4",  # rkm 81
             "CRA")  # rkm 94 (from SF Clearwater mouth)

# sf of sites of interest
sf_sites_sf = config %>%
  filter(site_code %in% sf_sites) %>%
  select(site_code, rkm, rkm_total, latitude, longitude) %>%
  distinct() %>%
  st_as_sf(coords = c("longitude", 
                      "latitude"), 
           crs = 4326)

# download subset of NHDPlus flowlines
# nhd_list = queryFlowlines(sites_sf = sf_sites_sf,
#                           root_site_code = "SC1",
#                           min_strm_order = 3,
#                           dwnstrm_sites = F)
# save(nhd_list, file = here("data/derived_data/spatial/sf_flowlines.rda"))
load(here("data/derived_data/spatial/sf_flowlines.rda"))

# velocity barrier
barrier_pt = tribble(~site, ~latitude, ~longitude,
                     "Velocity Barrier", 45.797042, -115.782912) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

# create map
sf_map = ggplot() +
  # flowlines
  geom_sf(data = nhd_list$flowlines,
          aes(color = as.factor(StreamOrde),
              size = StreamOrde)) +
  scale_color_viridis_d(direction = -1,
                        option = "D",
                        name = "Stream\nOrder",
                        end = 0.8) +
  scale_size_continuous(range = c(0.2, 1.2),
                        guide = 'none') +
  # basin outline
  geom_sf(data = nhd_list$basin,
          fill = NA,
          lwd = 2) +
  # add sites with labels
  geom_sf(data = sf_sites_sf,
          size = 3,
          color = "black") +
  ggrepel::geom_label_repel(
    data = sf_sites_sf,
    aes(label = site_code, 
        geometry = geometry),
    size = 3,
    stat = "sf_coordinates",
    min.segment.length = 0,
    max.overlaps = 50) +
  # velocity barrier point
  geom_sf(data = barrier_pt,
          size = 4,
          color = "red") +
  theme_bw() +
  theme(axis.title = element_blank())
sf_map

```

```{r sites-table}
sf_sites_tbl = config %>%
  filter(site_code %in% sf_sites) %>%
  select(site_code,
         site_name,
         rkm,
         latitude,
         longitude) %>%
  distinct() %>%
  mutate(rkm = c(94, 1, 2, 60, 81)) %>%
  arrange(rkm) %>%
  rename(
    Site = site_code,
    Name = site_name,
    Latitude = latitude,
    Longitude = longitude
  ) %>%
  kable(booktabs = T,
        align = "clccc",
        caption = "Instream PIT Tag Detection sites used in the evaluation including site codes, names, and location.") %>%
  kable_styling(full_width = F,
                position = "center",
                bootstrap_options = c("striped", "condensed"))
sf_sites_tbl

```


# Evaluation

## Data

Our evaluation used information from PIT-tagged adult steelhead and Chinook salmon detected at any IPTDS in the South Fork Clearwater River during their upstream migration to spawn. First, we queried [DART](https://www.cbr.washington.edu/dart) to identify all PIT tags observed at Lower Granite Dam for steelhead and Chinook salmon, during spawn years 2022 and 2023, using the `queryObsDART()` function in the R package `PITcleanr`, and filtered those observations to identify any returning adult later observed at any IPTDS in the South Fork Clearwater River (Table \@ref(tab:sites-table)). Note that adult detection rates on the Lower Granite Dam adult fish ladder are essentially 100%, and thus, any adult returning to the South Fork Clearwater River is expected to not be "missed". We then used the PIT-tag lists of returning adults to the South Fork Clearwater River and queried their complete tag histories in [PTAGIS](https://www.ptagis.org/) which includes their mark and release information which will be important for subsequent analyses. The complete tag histories were then "compressed" and "filtered" using the `compress()` and `filterDetections()` in `PITcleanr`, respectively. For steelhead, we removed any detections after May 31 during the spawn year and for Chinook salmon after September 15 to avoid detections of any kelts, shed tags, etc. The compressed and filtered detections for steelhead and Chinook salmon for spawn years 2022 and 2023 are available for download [here](https://github.com/NPTfisheries/SF_Clearwater_Passage/blob/main/data/derived_data/sf_clearwater_filtered_detections.csv). The filtered detections were also converted to capture histories using the `buildCapHist()` function in `PITcleanr`. Finally, biological and origin (e.g., natural- versus hatchery-origin) for each tag were pulled from the Lower Granite Dam trapping database (LGTrappingDB). All data and code used in this evaluation are available at a GitHub site [here](https://github.com/NPTfisheries/SF_Clearwater_Passage).

## Conversion Rates

We estimated conversion rates to each IPTDS for Chinook salmon and steelhead in spawn years 2022 and 2023 (Table \@ref(tab:conv-rates)). The conversion rate is an estimate of the percentage of PIT-tagged adults that arrived at a given site of those that passed the previous site. For example, in SY2022, we estimated that 12 PIT-tagged steelhead passed SC4, which is 15.6% of the 77 PIT-tagged steelhead estimated to have passed SC3. Table \@ref(tab:conv-rates) provides the number of tags observed, detection efficiencies, expanded (estimated) tags and conversion rates by IPTDS site, species, and spawn year. 

```{r conv-rates}
# summarize tags, efficiencies, and conversion rates
conv_rate_df = site_eff %>%
  group_by(species, spawn_year) %>%
  mutate(conv_rate_pct = round(est_tags_at_node / lag(est_tags_at_node) * 100, 1)) %>%
  ungroup() %>%
  mutate(across(eff_est:eff_se, ~ round(. * 100, digits = 1)))

conv_rate_df %>%
  select(
    Species = species,
    `Spawn Year` = spawn_year,
    Site = node,
    `Obs Tags` = tags_at_node,
    `Detection Eff (%)` = eff_est,
    `Detection Eff SE` = eff_se,
    `Est Tags` = est_tags_at_node,
    `Est Tags SE` = est_tags_se,
    `Conversion (%)` = conv_rate_pct
  ) %>%
  kable(booktabs = T,
        align = "lcccccccc",
        caption = "Summary of tags observed at and estimated passing each site for steelhead and Chinook salmon, spawn years 2022 and 2023, including site efficiencies and conversion rates.") %>%
  kable_styling(full_width = F,
                position = "center",
                bootstrap_options = c("striped", "condensed"))

```

Conversion rates to sites SC3 and SC4, within each species, are similar in both spawn years 2022 and 2023 (Figure \@ref(fig:conv-rate-p)).

```{r conv-rate-p, fig.cap = "Conversion rates for Chinook salmon and steelhead to sites SC3 and SC4 on the South Fork Clearwater River. Conversion rate represents the percentage of tagged fish estimated to have arrived at the site of those that passed the previous site downstream."}
conv_rate_df %>%
  select(species, spawn_year, node, conv_rate_pct) %>%
  mutate(spawn_year = as.factor(spawn_year)) %>%
  filter(node %in% c("SC3", "SC4")) %>%
  ggplot(aes(x = node, y = conv_rate_pct, fill = spawn_year)) +
  geom_col(position = "dodge") +
  theme_classic() +
  labs(x = "Site",
       y = "Conversion Rate (%)",
       fill = "Spawn Year") +
  facet_wrap(~species)

```

## Length and Saltwater Age

We summarized conversion rates by fork length (mm) and saltwater age for those adults that were sampled and/or PIT-tagged at Lower Granite Dam (Tables \@ref(tab:length-tbl) and \@ref(tab:sw-age-tbl)). If the river section of interest was problematic to adult passage due to high water velocities, we might anticipate that conversion rates are higher for larger or older fish because of increased fitness and swimming capabilities. Sample sizes are low, but no noticeable trends were observed in conversion rates by length or saltwater age (Tables \@ref(tab:length-tbl) and \@ref(tab:sw-age-tbl)).

```{r length-tbl}
# length bins and labels
bins = seq(floor(min(sf_df$lgr_fl_mm, na.rm = T) / 50) * 50,
           ceiling(max(sf_df$lgr_fl_mm, na.rm = T) / 50) * 50,
           by = 50)
tmp = paste0(bins, "-", lead(bins))
bin_labels = tmp[1:(length(tmp) - 1)]

# passage by length
pass_by_fl = sf_df %>%
  select(species,
         spawn_year,
         tag_code,
         pass_sc3,
         success,
         lgr_fl_mm) %>%
  filter(!is.na(lgr_fl_mm)) %>%
  # get only fish that at least arrived at SC3
  filter(pass_sc3 == T | success == T) %>%
  mutate(fl_bin = cut(lgr_fl_mm,
                      breaks = bins,
                      labels = bin_labels)) %>%
  group_by(species, fl_bin) %>%
  summarize(n_tags = n(),
            success = sum(success)) %>%
  mutate(conv_rate = round((success / n_tags) * 100, 1))

pass_by_fl %>%
  rename(
    Species = species,
    `FL bin (mm)` = fl_bin,
    `Obs Tags` = n_tags,
    `Obs Tags Passing` = success,
    `Conversion (%)` = conv_rate
  ) %>%
  kable(booktabs = T,
        align = "ccccc",
        caption = "The number of tags observed arriving at SC3 (Obs Tags), and of those, the number and percentage observed arriving at SC4 (Obs Tags Passing), by fork length (mm) bin.") %>%
  kable_styling(full_width = F,
                position = "center",
                bootstrap_options = c("striped", "condensed"))

```

```{r sw-age-tbl}
# passage by saltwater age
pass_by_sw_age = sf_df %>%
  select(species,
         spawn_year,
         tag_code,
         pass_sc3,
         success,
         bio_scale_final_age) %>%
  # convert scale age to saltwater age
  mutate(sw_age = gsub(".*:","", bio_scale_final_age)) %>%
  mutate(sw_age = case_when(
    sw_age == "A" ~ NA,
    TRUE ~ sw_age)) %>%
  filter(!is.na(sw_age)) %>%
  # get only fish that at least arrived at SC3
  filter(pass_sc3 == T | success == T) %>%
  group_by(species, sw_age) %>%
  summarize(n_tags = n(),
            success = sum(success)) %>%
  mutate(conv_rate = round((success / n_tags) * 100, 1))

pass_by_sw_age %>%
  rename(
    Species = species,
    `Saltwater Age` = sw_age,
    `Obs Tags` = n_tags,
    `Obs Tags Passing` = success,
    `Conversion (%)` = conv_rate
  ) %>%
  kable(booktabs = T,
        align = "ccccc",
        caption = "The number of tags observed arriving at SC3 (Obs Tags), and of those, the number and percentage observed arriving at SC4 (Obs Tags Passing), by saltwater age.") %>%
  kable_styling(full_width = F,
                position = "center",
                bootstrap_options = c("striped", "condensed"))

```

We summarized length-at-saltwater age relationships for both Chinook salmon and steelhead for PIT-tagged adults returning to the South Fork Clearwater River in spawn years 2022 and 2023 (Figure \@ref(fig:fl-by-age)).

```{r fl-by-age, fig.cap = "Fork length by salwater age for PIT-tagged adult Chinook salmon and steelhead observed in the South Fork Clearwater River, spawn years 2022 and 2023."}
# plot length-by-age, just of interest
sf_df %>%
  select(species,
         lgr_fl_mm,
         bio_scale_final_age) %>%
  filter(!is.na(lgr_fl_mm), !is.na(bio_scale_final_age)) %>%
  mutate(sw_age = gsub(".*:","", bio_scale_final_age)) %>%
  mutate(sw_age = case_when(
    sw_age == "A" ~ NA,
    TRUE ~ sw_age)) %>%
  filter(!is.na(sw_age)) %>%
  ggplot(aes(x = sw_age,
             y = lgr_fl_mm)) +
  geom_boxplot() +
  theme_classic() +
  labs(x = "Saltwater Age",
       y = "FL (mm)") +
  facet_wrap(~species)

```

We also summarized the fork lengths of PIT-tagged adults that arrived at the SC4 site (TRUE) versus those observed at SC3, but did not arrive at SC4 (FALSE; Figure \@ref(fig:fl-by-success)). Again, we might hypothesize that adults arriving to SC4 are larger than those that do not if water velocities were problematic for smaller or younger fish.

```{r fl-by-success, fig.cap = "Fork length of PIT-tagged adult Chinook salmon and steelhead that arrived at SC4 (TRUE) versus those that passed SC3, but did not arrive to SC4 (FALSE)."}
# compare length of fish observed arriving at SC4, versus those that did not
sf_df %>%
  select(species,
         tag_code,
         lgr_fl_mm,
         pass_sc3,
         success) %>%
  # get only fish that at least arrived at SC3
  filter(pass_sc3 == T | success == T) %>%
  filter(!is.na(lgr_fl_mm)) %>%
  ggplot(aes(x = success,
             y = lgr_fl_mm)) +
  geom_boxplot() +
  theme_classic() +
  labs(x = "Successful Passage to SC4",
       y = "FL (mm)") +
  facet_wrap(~species)

```

## Flow

```{r SC4-water-level}
# passage by water level
pass_by_wl = sf_df %>%
  select(species,
         spawn_year,
         tag_code,
         pass_sc3,
         success,
         sc4_water_level_m) %>%
  # get only fish that at least arrived at SC3
  filter(pass_sc3 == T | success == T) %>%
  filter(!is.na(sc4_water_level_m))

# water level bins and labels
wl_bins = seq(floor(min(pass_by_wl$sc4_water_level_m, na.rm = T) / 0.2) * 0.2,
              ceiling(max(pass_by_wl$sc4_water_level_m, na.rm = T) / 0.2) * 0.2,
              by = 0.2)
tmp = paste0(wl_bins, "-", lead(wl_bins))
wl_bin_labels = tmp[1:(length(tmp) - 1)]

pass_by_wl2 = pass_by_wl %>%
  mutate(wl_bin = cut(sc4_water_level_m,
                      breaks = wl_bins,
                      labels = wl_bin_labels)) %>%
  group_by(species, wl_bin) %>%
  summarize(n_tags = n(),
            success = sum(success)) %>%
  mutate(conv_rate = round((success / n_tags) * 100, 1))

pass_by_wl2 %>%
  rename(
    Species = species,
    `Water Level (m)` = wl_bin,
    `Obs Tags` = n_tags,
    `Obs Tags Passing` = success,
    `Conversion (%)` = conv_rate
  ) %>%
  kable(booktabs = T,
        align = "ccccc",
        caption = "The number of tags observed arriving at SC3 (Obs Tags), and of those, the number and percentage observed arriving at SC4 (Obs Tags Passing), by water level (m) at SC4.") %>%
  kable_styling(full_width = F,
                position = "center",
                bootstrap_options = c("striped", "condensed"))

```

```{r wl-by-success, fig.cap = "Water level (m) at the SC4 environmental probe at the time the PIT-tagged Chinook salmon or steelhead was last detected at SC1, SC2, or SC3 for those adults that successfully migrated to SC4 (TRUE) versus those that migrated to SC3, but did not arrive at SC4 (FALSE)."}
sf_df %>%
  select(species,
         tag_code,
         sc4_water_level_m,
         pass_sc3,
         success) %>%
  # get only fish that at least arrived at SC3
  filter(pass_sc3 == T | success == T) %>%
  filter(!is.na(sc4_water_level_m)) %>%
  ggplot(aes(x = success,
             y = sc4_water_level_m)) +
  geom_boxplot() +
  theme_classic() +
  labs(x = "Successful Passage to SC4",
       y = "Water Level (m) at SC4") +
  facet_wrap(~species)

```

```{r cfs}
# cfs bins and labels
cfs_breaks = 500
cfs_bins = seq(floor(min(sf_df$daily_mean_cfs, na.rm = T) / cfs_breaks) * cfs_breaks,
               ceiling(max(sf_df$daily_mean_cfs, na.rm = T) / cfs_breaks) * cfs_breaks,
               by = cfs_breaks)
tmp = paste0(cfs_bins, "-", lead(cfs_bins))
cfs_bin_labels = tmp[1:(length(tmp) - 1)]

# passage by cfs at Stites
pass_by_cfs = sf_df %>%
  select(species,
         tag_code,
         daily_mean_cfs,
         pass_sc3,
         success) %>%
  # get only fish that at least arrived at SC3
  filter(pass_sc3 == T | success == T) %>%
  filter(!is.na(daily_mean_cfs)) %>%
  mutate(cfs_bin = cut(daily_mean_cfs,
                       breaks = cfs_bins,
                       labels = cfs_bin_labels)) %>%
  group_by(species, cfs_bin) %>%
  summarize(n_tags = n(),
            success = sum(success)) %>%
  mutate(conv_rate = round((success / n_tags) * 100, 1))

pass_by_cfs %>%
  rename(
    Species = species,
    `Discharge (cfs)` = cfs_bin,
    `Obs Tags` = n_tags,
    `Obs Tags Passing` = success,
    `Conversion (%)` = conv_rate
  ) %>%
  kable(booktabs = T,
        align = "ccccc",
        caption = "The number of tags observed arriving at SC3 (Obs Tags), and of those, the number and percentage observed arriving at SC4 (Obs Tags Passing), by cfs near Stites, ID (USGS gage 13338500).") %>%
  kable_styling(full_width = F,
                position = "center",
                bootstrap_options = c("striped", "condensed"))

```

## Release Group

```{r release-groups}
# release groups in dataset
sf_df %>%
  select(species,
         tag_code, 
         rel_site) %>%
  group_by(species, rel_site) %>%
  summarise(n = n()) %>%
  rename(Species = species,
         `Release Site` = rel_site) %>%
  kable(booktabs = T,
        align = "ccc",
        caption = "Release groups present in PIT-tagged adult Chinook salmon and steelhead observed in the South Fork Clearwater River, spawn years 2022 and 2023 combined.") %>%
  kable_styling(full_width = F,
                position = "center",
                bootstrap_options = c("striped", "condensed"))

```

```{r success-by-rg}
# passage by release group
pass_by_rg = sf_df %>%
  select(species,
         spawn_year,
         tag_code,
         pass_sc3,
         success,
         rel_site) %>%
  mutate(rel_group = case_when(
    rel_site == "CLWRSF" ~ "South Fork Clearwater River",
    rel_site == "MEAD2C" ~ "Meadow Creek",
    rel_site == "NEWSOC" ~ "Newsome Creek",
    TRUE ~ "Other")) %>%
  group_by(species, rel_group) %>%
  summarize(n_tags = n(),
            success = sum(success)) %>%
  mutate(conv_rate = round((success / n_tags) * 100, 1))

pass_by_rg %>%
  rename(
    Species = species,
    `Release Group` = rel_group,
    `Obs Tags` = n_tags,
    `Obs Tags Passing` = success,
    `Conversion (%)` = conv_rate
  ) %>%
  kable(booktabs = T,
        align = "clccc",
        caption = "The number of tags observed arriving at SC3 (Obs Tags), and of those, the number and percentage observed arriving at SC4 (Obs Tags Passing), by release group.") %>%
  kable_styling(full_width = F,
                position = "center",
                bootstrap_options = c("striped", "condensed"))

```


# Discussion

## Limitations

First Newsome Creek juvenile releases occurred in spring 2022; 1-ocean adults will return to SF Clearwater in spring 2024 and 2-ocean adults in spring 2025.


# Literature Cited

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

